NoTitle

### 2017-06-12 Aplas17/sections/Introduction.tex

Some better workarounds include delegation, renaming/exclusion in the trait model, and so on, yet they still have various drawbacks as we will discuss in Section ?. On the other hand, there is indeed some related work [] that approaches our expectation, that is, they allow two unintentionally conflicted method to coexist in the system. Among them C++ is a representative that accepts the triangle inheritance and resolves the ambiguity by static dispatch. However, when virtual methods are defined with unintentional conflicts, C++ has to throw (linker?) errors on them.
This is again unsatisfactory because dynamic dispatch is necessary under many circumstances for code reuse and extensibility.

Having tolerance for unintentional method conflicts does not mean to sacrifice extensibility, hence in contrast with static dispatch and dynamic dispatch, we propose a third approach, which looks like a combination of both, called \textit{hierarchical dispatch}. By hierarchical dispatch, the method binder will look at both the static type and the dynamic type of the receiver during runtime. When there are multiple branches that cause unintentional conflicts, the static type can specify one branch among them for unambiguity, and the dynamic type helps to find the most specific (the newest) implementation. In that case both unambiguity and extensibility are preserved. To present this idea, we introduce a formalized model in Section ? based on Featherweight Java, together with some theorems and proofs for type soundness. In the model we also propose \textit{partial overrides}, where a method refinement can be applied only to one branch.

>>>>

Some better workarounds include delegation, renaming/exclusion in the trait model, and so on, yet they still have various drawbacks as we will discuss in Section ?. On the other hand, there is indeed some related work [] that is approaching our expectation, that is, they allow two unintentionally conflicted method to coexist in the system. Among them C++ is a representative that accepts the triangle inheritance and resolves the ambiguity by static dispatch. However, when virtual methods are defined with unintentional conflicts, C++ has to throw (linker?) errors on them.
This is again unsatisfactory because dynamic dispatch is necessary under many circumstances for code reuse and extensibility.

Having tolerance for unintentional method conflicts does not mean to sacrifice extensibility, hence in contrast with static dispatch and dynamic dispatch, we propose a third approach, which looks like a combination of both, called \textit{hierarchical dispatch}. By hierarchical dispatch, the method binder will look at both the static type and the dynamic type of the receiver during runtime. When there are multiple branches that cause unintentional conflicts, the static type can specify one branch among them for unambiguity, and the dynamic type helps to find the most specific (the newest) implementation. In that case both unambiguity and extensibility are preserved. To present this idea, we introduce a formalized model in Section ? based on Featherweight Java, together with some theorems and proofs for type soundness. In the model we also propose \textit{partial overrides}, where a method refinement can be applied only to one branch. Briefly speaking, our model generalizes the trait model by providing support to the triangle inheritance.

### 2017-06-12 Aplas17/sections/Overview.tex

\section{A Running Example: DrawableDeck}

\red{H: with UML graphs the illustration could be better.}
This section illustrates the features of our \MIM{} model for resolving unintentional method
conflicts. As mentioned before, such a case arises when two inherited methods happen to have the
same signature, but with different semantics and functionalities. This could be quite troublesome
to programmers that use multiple inheritance. Below we illustrate with a running example called \lstinline|DrawableDeck|.
Note that we use Java-like syntax throughout the paper, and all types are defined with the keyword ``\lstinline|interface|'', which
supports multiple inheritance. Since \MIM{} is designed to be simple, we do not allow abstract methods, that is every method
is required to have a body for its implementation. In that case, interfaces can be directly instantiated by the keyword ``\lstinline|new|''
to create an object.

@@

interface Drawable {
  void draw() { // draws something on the screen
    JFrame frame = new JFrame("Canvas");
    frame.setVisible(true);
    ...
  }
}
\end{lstlisting}\vspace{3pt}
Note that both methods have \lstinline|void| return type (we will not formalize
\lstinline|void| in our calculus afterwards; here is only for illustration). In \lstinline|Deck|, \lstinline|draw()| tries to get the cards as a stack, pops
out the top card, and so on. While in \lstinline|Drawable|, \lstinline|draw()|
creates a blank canvas using \lstinline|JFrame|. Now, a programmer is designing a
card game with GUI. He may want to draw a deck on the screen, so he defines a drawable
deck using multiple inheritance:

@@

\paragraph{I. Delegation.} As an alternative to multiple inheritance, delegation can be used by
introducing two fields with \lstinline|Drawable| type and \lstinline|Deck| type, respectively. This avoids
method conflicts, nevertheless, delegation itself is too restricted in modularity, and meanwhile
introduces a lot of boilerplate.

@@

\paragraph{IV. Method exclusion like traits.} Same reason as above.

\paragraph{V. Method renaming like traits.} This is probably what people do in most cases, by simply renaming one to avoid conflicts.
It can indeed preserve both features, however, it is cumbersome in practice, as introducing new names can affect other code blocks.
Certainly this is a workaround, not a solution.\\

@@

DrawableLoggingDeck d = new DrawableLoggingDeck();
d.shuffleAndDraw(); // ambiguous draw()
\end{lstlisting}\vspace{3pt}
Since the dynamic type of the receiver is \lstinline|DrawableLoggingDeck|, calling \lstinline|shuffleAndDraw()| triggers the ambiguity. When \lstinline|shuffleAndDraw()| invokes \lstinline|draw()|, what we really want is \lstinline|LoggingDeck.draw()|, yet
neither static dispatch nor dynamic dispatch in languages like C++ does so.
 Therefore, we need to find another algorithm for method binding.

@@

On the other hand, \lstinline|((I) e).m()|, or simply \lstinline|e.m()| where \lstinline|e| has static type \lstinline|I|, behaves the same as \lstinline|e.I::m()| in our model. Such a dispatch make uses of both the static type and the dynamic type of the receiver. Intuitively, the static type specifies one branch of the method to avoid ambiguity, and the dynamic type finds the latest version on that branch. It may still introduce ambiguity when there are multiple paths from the static type to the dynamic type, and those paths cause conflicts. To prevent this, we disallow this kind of conflict to ensure unambiguity. That is to say, we do not allow two methods to override a same base method when they cause conflicts. This is natural, as they are just two versions of the same operation, hence it is no longer an ``unintentional'' conflict in that way.

>>>>

\section{A Running Example: DrawableDeck}~\label{sec:overview}
\includegraphics[height=3.3cm]{pics/DrawableLoggingDeck.pdf}
\red{H: with UML graphs the illustration could be better.}
This section illustrates the features of our \MIM{} model for resolving unintentional method
conflicts. As mentioned before, such a case arises when two inherited methods happen to have the
same signature, but with different semantics and functionalities. This could be quite troublesome
to programmers that use multiple inheritance. Below we illustrate with a running example called \lstinline|DrawableDeck|.
Note that we use Java-like syntax throughout the paper, and all types are defined with the keyword ``\lstinline|interface|'', which
supports multiple inheritance. Since \MIM{} is designed to be simple, we do not allow abstract methods, that is every method
is required to have a body for its implementation. In that case, interfaces can be directly instantiated by the keyword ``\lstinline|new|''
to create an object. \haoyuan{TODO: mention relation to Java 8 interfaces with default methods; and traits}

@@

interface Drawable {
  void draw() { // draws something on the screen
    JFrame frame = new JFrame("Canvas");
    frame.setVisible(true);
    ...
  }
}
\end{lstlisting}\vspace{3pt} \haoyuan{TODO: drawObject}
Note that both methods have \lstinline|void| return type (we will not formalize
\lstinline|void| in our calculus afterwards; here is only for illustration). In \lstinline|Deck|, \lstinline|draw()| tries to get the cards as a stack, pops
out the top card, and so on. While in \lstinline|Drawable|, \lstinline|draw()|
creates a blank canvas using \lstinline|JFrame|. Now, a programmer is designing a
card game with GUI. He may want to draw a deck on the screen, so he defines a drawable
deck using multiple inheritance:

@@

\paragraph{I. Delegation.} As an alternative to multiple inheritance, delegation can be used by
introducing two fields with \lstinline|Drawable| type and \lstinline|Deck| type, respectively. This avoids
method conflicts, nevertheless, delegation itself is too restricted in modularity, and meanwhile
introduces a lot of boilerplate.\haoyuan{TODO: drawbacks: hard to correctly maintain self-references[]}

@@

\paragraph{IV. Method exclusion like traits.} Same reason as above. \haoyuan{TODO: put exclusion and renaming together}

\paragraph{V. Method renaming like traits.} This is probably what people do in most cases, \haoyuan{not supported by most languages; breaks subtyping} by simply renaming one to avoid conflicts.
It can indeed preserve both features, however, it is cumbersome in practice, as introducing new names can affect other code blocks.
Certainly this is a workaround, not a solution.\\

@@

DrawableLoggingDeck d = new DrawableLoggingDeck();
d.shuffleAndDraw(); // ambiguous draw()
\end{lstlisting}\vspace{3pt}
Since the dynamic type of the receiver is \lstinline|DrawableLoggingDeck|, calling \lstinline|shuffleAndDraw()| triggers the ambiguity. When \lstinline|shuffleAndDraw()| invokes \lstinline|draw()|, what we really want is \lstinline|LoggingDeck.draw()|, yet
neither static dispatch nor dynamic dispatch in languages like C++ does so.
 Therefore, we need to find another algorithm for method binding\haoyuan{method resolution}.

@@

On the other hand, \lstinline|((I) e).m()| \haoyuan{explicit casts are not supported in our model}, or simply \lstinline|e.m()| where \lstinline|e| has static type \lstinline|I|, behaves the same as \lstinline|e.I::m()| in our model. Such a dispatch make uses of both the static type and the dynamic type of the receiver. Intuitively, the static type specifies one branch of the method to avoid ambiguity, and the dynamic type finds the latest version on that branch. It may still introduce ambiguity when there are multiple paths from the static type to the dynamic type, and those paths cause conflicts. To prevent this, we disallow this kind of conflict to ensure unambiguity. That is to say, we do not allow two methods to override a same base method when they cause conflicts. This is natural, as they are just two versions of the same operation, hence it is no longer an ``unintentional'' conflict in that way.
